Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-08-08 slides/performanceModels.tex
presented slides

\date{\includegraphics[width=.25\linewidth]{icpp.png}\\August 8, 2019}
\newcommand{\highlight}[1]{\colorbox{red}{$\displaystyle #1$}}
\pdfcompresslevel=2             %-- 0 = none, 9 = best

@@

\logo{%
	\makebox[0.95\paperwidth]{%
		\includegraphics[width=.3\linewidth]{logo_pnnl.png}
		\hfill%
		\includegraphics[width=.3\linewidth]{logo_inria.png}
	}%
}

@@

\begin{frame}{Task Graphs and Runtime Systems}
%%   \begin{itemize}
%%   	 \item Applications can be expressed as task graphs
%%\end{itemize}
		\begin{columns}
	\null \hfill
	\begin{column}{0.4\linewidth}
		\begin{center}
			\includegraphics[scale=0.2]{diagrams/taskGraph.eps}
		\end{center}
	\end{column}
	\begin{column}{0.5\linewidth}
		\begin{itemize}
			   	 \item Applications can be expressed as task graphs
			   	 \item Vertices represent tasks
			   	 \item Edges represent dependencies among tasks
		\end{itemize}
	\end{column}
\end{columns}
%%		\begin{center}
%%	\includegraphics[scale=0.2]{diagrams/taskGraph.eps}
%%\end{center}
\begin{block}{Task Based Runtime Systems}
		\begin{itemize}
	\item StarPU, QUARK, PaRSEC, StarSs, Legion, KAAPI
%%	\item Manages scheduling of computations and communications
	\item May only see a set of ready (independent) tasks
	\item A memory node may require data from other memory nodes 
%%	\item Output data is managed by runtime based on demand

	\only<1>{ \item Order of data transfers such that communication-computation overlap is maximized}
	\only<2>{\item Order of data transfers \textcolor{red}{between two memory nodes} such that communication-computation overlap is maximized}
\end{itemize}
\end{block}

\section{Problem Definition}
\begin{frame}{Problem Definition}
\begin{itemize}
	\item Communication and computation times are known in advance
	\begin{itemize}
		\item Approximated based on number of computations and hardware details
		\item Obtained from previous executions
	\end{itemize}
%%\item Tasks do not produce output data (or store in a separate buffer)
\end{itemize}
\begin{block}{Problem $DT$ }
	\begin{itemize}
		\item A set of tasks $ST=\{T_1, \cdots, T_n\}$ is scheduled on a processing unit $P$ with
		memory unit $M$ of capacity $C$
		\item Input data for tasks of $ST$
		reside on another memory unit
		\item Tasks do not produce output data
		\item Tasks do not require intermediate memory
		\item A tasks uses an amount of memory in $M$ from the
		start of its communication to the end of its computation
	\end{itemize}
\noindent Given $L$, is there a feasible schedule $S$ for $ST$ such that
makespan of $S$, $\mu(S) \le L$?
\end{block}

@@

\begin{frame}{Approach for the optimality proof}
\vspace*{-0.5cm}
\begin{center}
\includegraphics[scale=0.32]{./diagrams/original_swapped_schedules-eps-converted-to.pdf}
\end{center}

@@

\subsection{Limited Memory Capacity}
\begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsubsection]
\end{frame}
\newcommand{\threepart}{\textsc{3Par}\xspace}
\begin{frame}{Limited Memory Capacity}
\begin{itemize}
	\item Memory is required only to store input data (from the definition of our problem)
	\item As Communication time $\propto$ amount of communication, for each task:
	\begin{itemize}
		\item Communication time = Amount of communication  (\textcolor{green}{for simplification})
		\item Communication time = Amount of input data
	\end{itemize}
\end{itemize}

@@

	\begin{column}[c]{.35\linewidth}
		\footnotesize
		\begin{center}
\begin{tabular}{|c|c|c|c|}
	\hline
	Task & Memory & Comm & Comp\\
	& Req & Time & Time\\ \hline 
	%%				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
	%%				&=Comm Volume && \\ \hline
	A & 0 & 0 & 5\\ \hline
	B & 4 & 4 & 3\\ \hline
	C & 1 & 1 & 6\\ \hline
	D & 3 & 3 & 7\\ \hline
	E & 6 & 6 & 0.5\\ \hline
	F & 7 & 7 & 0.5\\ \hline
\end{tabular}
		\end{center}
	\end{column}

@@

\begin{frame}{Static Order with Dynamic Corrections Based Strategies}
\begin{columns}
	\footnotesize
	\begin{column}[c]{0.4\linewidth}
		\begin{tabular}{|c|c|c|c|}
			\hline
	Task & Memory & Comm & Comp\\
& Req & Time & Time\\ \hline 
			%%			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
			%%			&=Comm Volume && \\ \hline
			A & 4 & 4 &  1\\ \hline
			B & 2 & 2 & 6\\ \hline
			C & 8 & 8 & 8\\ \hline
			D & 5 & 5 & 4\\ \hline
			E & 3 & 3 & 2\\ \hline
		\end{tabular}
	\end{column}
	\begin{column}[c]{0.6\linewidth}
		\begin{block}{Memory Capacity=9}
%%\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
%%\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
%%\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
%%\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
%%\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
\newcommand{\taskA}[2][0]{\task[####1]{$A$}{####2}{4}{1}{cyan}}
\newcommand{\taskB}[2][0]{\task[####1]{$B$}{####2}{2}{6}{cyan!50!black}}
\newcommand{\taskC}[2][0]{\task[####1]{$C$}{####2}{8}{8}{blue!40!white}}
\newcommand{\taskD}[2][0]{\task[####1]{$D$}{####2}{5}{4}{blue!70!white}}
\newcommand{\taskE}[2][0]{\task[####1]{$E$}{####2}{3}{2}{blue}}
\begin{tikzpicture}[yscale=0.6,xscale=0.18]
\begin{scope}
\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
\taskB{0}
\taskD[1]{2}
\taskA{8}
\taskE{12}
\taskC{17}
\end{scope}
\begin{scope}[yshift=-2.75cm]
\schedule{36}{OOSCMR}{2,5,8,11,15,19,27,35}
\taskB{0}
\taskE[3]{2}
\taskA[1]{5}
\taskD{10}
\taskC{19}
\end{scope}
\begin{scope}[yshift=-5.5cm]
\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
\taskB{0}
\taskD[1]{2}
\taskE[1]{8}
\taskA{12}
\taskC{17}
\end{scope}
\end{tikzpicture}
		\end{block}
	\end{column}	
\end{columns}
\end{frame}

@@

\begin{block}{\textit{Bin Packing} ($BP$)}
	\begin{itemize}
		\item First-Fit algorithm for the bin packing problem
		\item Tasks added to the first bin in which they fit
		\item If no bin is found then a new bin is created
		\item Sequence made of all tasks from consecutive bins
	\end{itemize}
\end{block}
\end{frame}

@@

\begin{frame}{CCSD performance with $m_c=1.8GB$}
\begin{center}
	\includegraphics[width=1.05\linewidth]{./diagrams/results/ratio_to_optimal_selected_ccsd.pdf}
\end{center}
$\bullet$ Highly heterogeneous tasks\\
$\bullet$ Static order with dynamic correction variants outperform others in most cases
\end{frame}

\begin{frame}{Best variants of all categories}
%%\begin{columns}
%%	\begin{column}{0.55\linewidth}
\begin{block}{\only<1>{HF} \only<2>{CCSD} Performance}
	\begin{center}
\only<1>{\includegraphics[width=0.85\linewidth]{./diagrams/results/ratio_to_optimal_hf-best.pdf}}
\only<2>{\includegraphics[width=0.85\linewidth]{./diagrams/results/ratio_to_optimal_ccsd-best.pdf}}
	\end{center}
\end{block}
\end{frame}

\begin{frame}{Implementation challenges}
\begin{itemize}
	\vfill
	\item Impact of Congestion on communication/computation times
	\vfill
	\item Output data of a task
	\vfill
	\item Routes between two memory nodes
	\vfill
	\item Communications from multiple memory nodes can happen at the same time
	\vfill
\end{itemize}
\end{frame}

 \section{Conclusion and Ongoing Work}
 \begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsection]
\end{frame}
 \begin{frame}{Conclusion and Ongoing Work}
Conclusion:
 \begin{itemize}
  \item Problem of determining the optimal order of data transfers is NP-complete
  \item Our heuristics achieve significant overlap for HF and CCSD applications
 \end{itemize}
\vfill
Ongoing work: 
%% \end{frame}
%% \begin{frame}{Future work}
 \begin{itemize}
  \item Evaluation of our heuristics in the context of accelerators
  \item Implementation of the proposed heuristics
 \item Automatic selection of the best heuristic
 \item Model bandwidth sharing to support multiple simultaneous communications
 \end{itemize}
 \end{frame}

%% \begin{frame}
%%  \begin{center}
%%   {\huge Thank \newline You!}
%%    
%%  \end{center}
%% \end{frame}

>>>>

\date{\includegraphics[width=.25\linewidth]{./diagrams/icpp.png}\\August 8, 2019\\ Kyoto, Japan}
\newcommand{\highlight}[1]{\colorbox{red}{$\displaystyle #1$}}
\pdfcompresslevel=2             %-- 0 = none, 9 = best

@@

\logo{%
	\makebox[0.95\paperwidth]{%
		\includegraphics[width=.3\linewidth]{./logo/logo_pnnl.png}
		\hfill%
		\includegraphics[width=.3\linewidth]{./logo/logo_inria.png}
	}%
}

@@

\begin{frame}{Task Graphs and Runtime Systems}
%%   \begin{itemize}
%%   	 \item Applications can be expressed as task graphs
%%\end{itemize}
		\begin{columns}
	\null \hfill
	\begin{column}{0.4\linewidth}
		\begin{center}
			\includegraphics[scale=0.2]{diagrams/taskGraph.eps}
		\end{center}
	\end{column}
	\begin{column}{0.5\linewidth}
		\begin{itemize}
			   	 \item Application can be expressed as a task graph
			   	 \item Vertices represent tasks
			   	 \item Edges represent dependencies among tasks
		\end{itemize}
	\end{column}
\end{columns}
%%		\begin{center}
%%	\includegraphics[scale=0.2]{diagrams/taskGraph.eps}
%%\end{center}
\begin{block}{Task Based Runtime Systems}
		\begin{itemize}
	\item QUARK, PaRSEC, StarPU, StarSs, Legion
%%	\item Manages scheduling of computations and communications
	\item May only see a set of ready (independent) tasks
	\item A memory node may require data from other memory nodes 
%%	\item Output data is managed by runtime based on demand

	\only<1>{ \item Order of data transfers such that communication-computation overlap is maximized}
	\only<2>{\item Order of data transfers \textcolor{red}{between two memory nodes} such that communication-computation overlap is maximized}
\end{itemize}
\end{block}
\end{frame}

\section{Problem Definition}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsection, subsectionstyle=shaded]
\end{frame}

\begin{frame}{Problem Definition}
\begin{itemize}
	\item Communication and computation times are known
	\begin{itemize}
		\item Based on number of computations and hardware details
		\item Obtained from previous executions (or iterations)
	\end{itemize}
%%\item Tasks do not produce output data (or store in a separate buffer)
\end{itemize}
\begin{block}{Problem $DT$ }
	\begin{itemize}
		\item A set of tasks $ST=\{T_1, \cdots, T_n\}$ is scheduled on a processing unit $P$ with
		memory unit $M$ of capacity $C$
		\item Input data for tasks of $ST$
		reside on another memory unit
		\item Tasks do not produce output data
		\item Tasks do not require intermediate memory
		\item A tasks uses an amount of memory in $M$ from the
		start of its communication to the end of its computation
	\end{itemize}
\noindent Given $L$, is there a feasible schedule $S$ for $ST$ such that
makespan of $S$, $\mu(S) \le L$?
\end{block}

@@

\begin{frame}{Approach for the Optimality Proof}
\vspace*{-0.5cm}
\begin{center}
\includegraphics[scale=0.32]{./diagrams/original_swapped_schedules-eps-converted-to.pdf}
\end{center}

@@

\subsection{Limited Memory Capacity}
\begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsubsection]
\end{frame}
\newcommand{\threepart}{\textsc{3Par}\xspace}
\begin{frame}{Limited Memory Capacity}
\begin{itemize}
	\item Memory is required only to store input data (from the definition of our problem)
	\item As communication time $\propto$ amount of communication, for each task:
	\begin{itemize}
		\item Communication time = Amount of communication  (\textcolor{green}{for simplification})
		\item Communication time = Amount of input data
	\end{itemize}
\end{itemize}

@@

	\begin{column}[c]{.35\linewidth}
		\footnotesize
		\begin{center}
\begin{tabular}{|c|c|c|c|}
	\hline
	Task & Memory & Comm & Comp\\
	& Req & Time & Time\\ \hline 
	%%				\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
	%%				&=Comm Volume && \\ \hline
	A & 0 & 0 & 5\\ \hline
	B & 4 & 4 & 3\\ \hline
	C & 1 & 1 & 6\\ \hline
	D & 3 & 3 & 7\\ \hline
	E & 6 & 6 & 0.5\\ \hline
	F & 7 & 7 & 0.5\\ \hline
\end{tabular}
		\end{center}

Memory Capacity = $10$
	\end{column}

@@

\begin{frame}{Static Order with Dynamic Corrections Based Strategies}
\begin{columns}
	\footnotesize
	\begin{column}[c]{0.4\linewidth}
		\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline
	Task & Memory & Comm & Comp\\
& Req & Time & Time\\ \hline 
			%%			\multirow{2}{*}{Task} & Memory Req & \multirow{2}{*}{Comm Time} & \multirow{2}{*}{Comp Time}\\  
			%%			&=Comm Volume && \\ \hline
			A & 4 & 4 &  1\\ \hline
			B & 2 & 2 & 6\\ \hline
			C & 8 & 8 & 8\\ \hline
			D & 5 & 5 & 4\\ \hline
			E & 3 & 3 & 2\\ \hline
		\end{tabular}
\end{center}

$OMIM$ order = $BCDEA$
	\end{column}
	\begin{column}[c]{0.6\linewidth}
		\begin{block}{Memory Capacity=9}
%%\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
%%\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
%%\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
%%\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
%%\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
\newcommand{\taskA}[2][0]{\task[####1]{$A$}{####2}{4}{1}{cyan}}
\newcommand{\taskB}[2][0]{\task[####1]{$B$}{####2}{2}{6}{cyan!50!black}}
\newcommand{\taskC}[2][0]{\task[####1]{$C$}{####2}{8}{8}{blue!40!white}}
\newcommand{\taskD}[2][0]{\task[####1]{$D$}{####2}{5}{4}{blue!70!white}}
\newcommand{\taskE}[2][0]{\task[####1]{$E$}{####2}{3}{2}{blue}}
\begin{tikzpicture}[yscale=0.6,xscale=0.18]
\begin{scope}
\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
\taskB{0}
\taskD[1]{2}
\taskA{8}
\taskE{12}
\taskC{17}
\end{scope}
\begin{scope}[yshift=-2.75cm]
\schedule{36}{OOSCMR}{2,5,8,11,15,19,27,35}
\taskB{0}
\taskE[3]{2}
\taskA[1]{5}
\taskD{10}
\taskC{19}
\end{scope}
\begin{scope}[yshift=-5.5cm]
\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
\taskB{0}
\taskD[1]{2}
\taskE[1]{8}
\taskA{12}
\taskC{17}
\end{scope}
\end{tikzpicture}
		\end{block}
	\end{column}	
\end{columns}
\end{frame}

@@

\begin{block}{\textit{Bin Packing} ($BP$)}
	\begin{itemize}
		\item First-Fit algorithm for the bin packing problem
		\item Tasks added to the first bin in which they fit
		\item If no bin is found then a new bin is created
		\item Sequence is made of all tasks from consecutive bins
	\end{itemize}
\end{block}
\end{frame}

@@

\begin{frame}{CCSD Performance with $m_c=1.8GB$}
\begin{center}
	\includegraphics[width=1.05\linewidth]{./diagrams/results/ratio_to_optimal_selected_ccsd.pdf}
\end{center}
$\bullet$ Highly heterogeneous tasks\\
$\bullet$ Static order with dynamic correction variants outperform others in most cases
\end{frame}

\begin{frame}{Best Variants of All Categories}
%%\begin{columns}
%%	\begin{column}{0.55\linewidth}
\begin{block}{\only<1>{HF} \only<2>{CCSD} Performance}
	\begin{center}
\only<1>{\includegraphics[width=0.85\linewidth]{./diagrams/results/ratio_to_optimal_hf-best.pdf}}
\only<2>{\includegraphics[width=0.85\linewidth]{./diagrams/results/ratio_to_optimal_ccsd-best.pdf}}
	\end{center}
\end{block}
\end{frame}

\begin{frame}{Implementation Challenges}
\begin{itemize}
	\vfill
	\item Impact of congestion on communication/computation times
	\vfill
	\item Output data of a task
	\vfill
	\item Multiple routes between two memory nodes
	\vfill
	\item Communications from multiple memory nodes can happen at the same time
	\vfill
\end{itemize}
\end{frame}

 \section{Conclusion and Ongoing Work}
 \begin{frame}
\frametitle{Table of Contents}
\tableofcontents[currentsection]
\end{frame}
 \begin{frame}{Conclusion and Ongoing Work}
Conclusion:
 \begin{itemize}
  \item Problem of determining the optimal order of data transfers is NP-complete
  \item Our heuristics achieve significant overlap for HF and CCSD applications
 \end{itemize}
\vfill
Ongoing work: 
%% \end{frame}
%% \begin{frame}{Future work}
 \begin{itemize}
  \item Evaluation of our heuristics in the context of accelerators
  \item Implementation of the proposed heuristics
 \item Automatic selection of the best heuristic
%% \item Model bandwidth sharing to support multiple simultaneous communications
 \end{itemize}
 \end{frame}

 \begin{frame}
  \begin{center}
   {\huge Thank \newline You!}

  \end{center}
 \end{frame}
